The Impact of Anthropogenic Forcing on ENSO Amplitude

### 2022-03-02 feedback/presentation_main.tex
stuff!

% Created 2022-01-06 Thu 19:00
% Intended LaTeX compiler: pdflatex
\documentclass[little]{basic}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\setcounter{secnumdepth}{1}
\author{Benjamin Goldman}
\date{\today}
\title{Feedback for main presentation}
\hypersetup{
 pdfauthor={Benjamin Goldman},
 pdftitle={Feedback for main presentation},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 27.2 (Org mode 9.5)}, 
 pdflang={English}}
\begin{document}

@@

\section{Feedback 1}
\label{sec:orgb0ce4b1}
Date: \textit{<2021-09-26 Sun>}
Time: 20:20
Listener: Abby Goldman
\subsection*{Feedback}
\label{sec:orgeae9fd0}
\begin{itemize}
\item lots of arm waving. SOOOOMUCH ARM WAVING!!! And very little breathing
\item make a list of “overhead” - make sure all concepts are well defined, in your notes, possibly in your slides
\item summary of what your many slides are — many different crunching of different data to circle around a process that we’re trying to under
\item insert heading slides, overall need a good deal more organization.
\item slide 1: break up into two or even 3?
\item Slide 2: Need to explain more generally what’s in the figure
\item Slide 3: climate simulation — is this method? If so, say so
\item Slide 4: ENSO and climate change slide — have “literature review” in heading.
\item slide about Mentor: do a better job explaining the interactive process between you and your mentor.
\item 3.4 index - whahah? I didn’t follow
\item Single forcing ensembles slide — not sure if all those arrows are working for you. Skip over single forcing setup, maybe?
\item use consistent language about the aerosols and their influence, considering that it’s a prediction
\end{itemize}

\section{Feedback 2}
\label{sec:org3a9f9f0}
Date: \textit{<2021-09-30 Thu>}
Time: 8:20
Listener: Marcus Chung, Sean O'Donnell, Jimena Perez
\subsection*{Feedback:}
\label{sec:orgd187f3b}
\begin{itemize}
\item Figure text is too small
\item Caption text can be smaller to make room
\item Say some transition for the conclusion heading slide
\item Clarify difference between El Niño and La Niña
\begin{itemize}
\item overall make intro 2 more organized
\end{itemize}
\item Good understanding and well communicated
\item Are some slides too empty-looking?
\item Say fewer words because it is over time and talking too fast
\item Use note cards
\item Periods should be consistent.
\end{itemize}

\section{Feedback 3}
\label{sec:org304f650}
Date: \textit{<2021-09-30 Thu>}
Time: 16:00
Listener: Michael Goldman
\subsection*{Feedback:}
\label{sec:org2526349}
\begin{itemize}
\item Make a few bullets clearer
\item Good tone
\item Slow down pacing
\item Introduction is slightly unclear
\item Use note cards
\end{itemize}

\section{Feedback 4}
\label{sec:org661561a}
Date: \textit{<2021-10-17 Sun>}
Time: 17:12
Listener: Jimena Perez
\subsection*{Feedback:}
\label{sec:orgfb0e433}
\begin{itemize}
\item Make figures bigger and make captions smaller
\item Love how detailed the content is
\item Great application slide
\item Make sure you explain your results in the context of your study, think big picture
\item Intro slide 2 is great we’ll explained
\end{itemize}

\section{Feedback 5}
\label{sec:org4ee2cd0}

Date: \textit{<2022-01-06 Thu>}
Time: 09:20
Listener: Jimena Perez
\subsection*{Feedback:}
\label{sec:orgcae29f9}
\begin{itemize}
\item Put Student vs mentor role earlier in presentation
\item Slow down speech
\item Write a script for wavelet slide
\item Don't admit you don't understand it
\item No monospace for urls
\item Clean up global warming slide
\end{itemize}
\end{document}

>>>>

% Created 2022-01-27 Thu 09:32
% Intended LaTeX compiler: pdflatex
\documentclass[little]{basic}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\setcounter{secnumdepth}{1}
\author{Benjamin Goldman}
\date{\today}
\title{Feedback for main presentation}
\hypersetup{
 pdfauthor={Benjamin Goldman},
 pdftitle={Feedback for main presentation},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 27.2 (Org mode 9.5)}, 
 pdflang={English}}
\begin{document}

@@

\section{Feedback 1}
\label{sec:orge5d83f9}
Date: \textit{<2021-09-26 Sun>}
Time: 20:20
Listener: Abby Goldman
\subsection*{Feedback}
\label{sec:org59ea01a}
\begin{itemize}
\item lots of arm waving. SOOOOMUCH ARM WAVING!!! And very little breathing
\item make a list of “overhead” - make sure all concepts are well defined, in your notes, possibly in your slides
\item summary of what your many slides are — many different crunching of different data to circle around a process that we’re trying to under
\item insert heading slides, overall need a good deal more organization.
\item slide 1: break up into two or even 3?
\item Slide 2: Need to explain more generally what’s in the figure
\item Slide 3: climate simulation — is this method? If so, say so
\item Slide 4: ENSO and climate change slide — have “literature review” in heading.
\item slide about Mentor: do a better job explaining the interactive process between you and your mentor.
\item 3.4 index - whahah? I didn’t follow
\item Single forcing ensembles slide — not sure if all those arrows are working for you. Skip over single forcing setup, maybe?
\item use consistent language about the aerosols and their influence, considering that it’s a prediction
\end{itemize}

\section{Feedback 2}
\label{sec:orgf550f9c}
Date: \textit{<2021-09-30 Thu>}
Time: 8:20
Listener: Marcus Chung, Sean O'Donnell, Jimena Perez
\subsection*{Feedback:}
\label{sec:org46618e2}
\begin{itemize}
\item Figure text is too small
\item Caption text can be smaller to make room
\item Say some transition for the conclusion heading slide
\item Clarify difference between El Niño and La Niña
\begin{itemize}
\item overall make intro 2 more organized
\end{itemize}
\item Good understanding and well communicated
\item Are some slides too empty-looking?
\item Say fewer words because it is over time and talking too fast
\item Use note cards
\item Periods should be consistent.
\end{itemize}

\section{Feedback 3}
\label{sec:org4fbc4a2}
Date: \textit{<2021-09-30 Thu>}
Time: 16:00
Listener: Michael Goldman
\subsection*{Feedback:}
\label{sec:orgb1ba02d}
\begin{itemize}
\item Make a few bullets clearer
\item Good tone
\item Slow down pacing
\item Introduction is slightly unclear
\item Use note cards
\end{itemize}

\section{Feedback 4}
\label{sec:org94d9adb}
Date: \textit{<2021-10-17 Sun>}
Time: 17:12
Listener: Jimena Perez
\subsection*{Feedback:}
\label{sec:orgff9e78c}
\begin{itemize}
\item Make figures bigger and make captions smaller
\item Love how detailed the content is
\item Great application slide
\item Make sure you explain your results in the context of your study, think big picture
\item Intro slide 2 is great we’ll explained
\end{itemize}

\section{Feedback 5}
\label{sec:org71846e3}

Date: \textit{<2022-01-06 Thu>}
Time: 09:20
Listener: Jimena Perez
\subsection*{Feedback:}
\label{sec:orga0d61f9}
\begin{itemize}
\item Put Student vs mentor role earlier in presentation
\item Slow down speech
\item Write a script for wavelet slide
\item Don't admit you don't understand it
\item No monospace for urls
\item Clean up global warming slide
\end{itemize}

\section{Feedback 6}
\label{sec:orge523970}

Date: \textit{<2022-01-15 Sat>}
Time: 09:20
Listener: Abby Goldman
\subsection*{Feedback}
\label{sec:org2833856}
\begin{itemize}
\item Make some notes to follow
\item Couple slides were unclear
\item Speak slower and less nervous
\item Good introduction
\end{itemize}

\section{Feedback 7}
\label{sec:org3149d17}
Date: \textit{<2022-01-20 Thu>}
Time: 9:25
Listener: Jimena Perez
\subsection*{Feedback:}
\label{sec:org43a5e7b}
\begin{itemize}
\item Love how the intro slides flow, the flow chart in the methodology slide becomes easier to understand every time you present it
\item Maintain speaking pace, describe implications of methodology while you're going through methodolgy.
\end{itemize}

\section{Feedback 8}
\label{sec:org6818510}
Date: \textit{<2022-01-25 Tue>}
Time: 9:25
Listener: Sean O'Donnell
\subsection*{Feedback:}
\label{sec:orge3ca3f6}
\begin{itemize}
\item Don't have to say ziggy zaggy line on 2nd slide (say red line or blue line)
\item Don't think you described figure 2
\item Liked explanation of figure 3 and El Nino/La nina in general
\item Yes, def mention ENSO first before EL Nino and La nina in introduction
\item Liked simplicity of powerpoint design, made it very easy to follow
\item Maybe mention CESM before stating research questions so people can understand its background (may not be necessary though)?
\item Strong explanation of methodology (liked the flowchart)
\item Liked the addition of ``increasing variance'' and decreasing variance`` to Figure 4
\item What's difference between CESM1 and CESM 2 (may have missed this in presentation)?
\item Yes, def note software used before acknowledgements (don't have to describe each database)
\end{itemize}
\end{document}

### 2022-03-02 presentations/presentation/presentation.tex
Presentation is 16:10, not 4:3
stuff!

\documentclass{beamer}

@@

\begin{frame}{Climate Change: Global Warming vs. Climate Variability}
  \begin{figure}
    \centering
    \includegraphics[width=0.8\textwidth]{figures/intro_fig_3.pdf}
    \caption{Global mean land air temperature in GISSTEMP 4 dataset. \citep{gistemp2019giss} and \citep{lenssen2019improvements}}
    \label{fig:woohoo}
  \end{figure}
\end{frame}

\begin{frame}{Greenhouse Gasses and Friends}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
    \item \alert{Forcing}: any external factor that affects climate.
      \begin{description}
      \item[\alert{GHG}] Greenhouse gasses
      \item[\alert{AER}] Aerosols (natural: volcanic ash, artificial: smoke)
      \item[\alert{BMB}] Biomass burning
      \item[\alert{LULC}] Land use/cover (deforestation, desertification)
      \end{description}
    \end{itemize}
    \column{0.6\textwidth}
    \myfig{greenhouse_Effect.jpg}{Factors that contribute to the greenhouse effect. \url{https://www.coolaustralia.org/the-greenhouse-effect-secondary}}{this}
  \end{columns}
\end{frame}

\begin{frame}{El Niño, La Niña, and ENSO}
  \begin{figure}
    \begin{columns}
      \column{0.5\textwidth}
      \begin{itemize}
      \item ENSO = El Niño/Southern Oscillation
      \item El Niño is the warm phase, and La Niña is the cool phase.
      \item Warming and cooling of the Pacific Ocean.
      \item Affects human societies through temperature and rainfall. \citep{ropelewski1987global}
      \end{itemize}
      \caption{Comparison of SST anomaly between 1975 La Niña event and 1997 El Niño event in HadISST 1 dataset. \citep{rayner2003global}}
      \column{0.6\textwidth}
      \includegraphics[width = \textwidth]{figures/intro_fig.pdf}
    \end{columns}
  \end{figure}
\end{frame}

@@

\begin{frame}{Gap and Questions}
  \alert{Gap}
  \begin{itemize}
  \item Little research using a large ensemble to examine the effect of individual factors on ENSO.
  \item Considerable disagreement between studies on whether ENSO will strengthen or weaken due to global warming
  \end{itemize}
  \alert{Questions}
  \begin{enumerate}
  \item Do the CESM1 and CESM2 predict increased or decreased ENSO intensity in the future?
  \item Is the predicted increase (or decrease) due to human activities?
  \end{enumerate}
\end{frame}

@@

\begin{frame}{Data: Community Earth System Model}
  \begin{itemize}
  \item Community Earth System Model (CESM) Versions 1 and 2 \citep{kay2015community} \citep{danabasoglu2020community}.
  \item Predicts climate over 21st century with global warming.
  \item Ensemble: collection of multiple simulations.
  \item Single forcing ensembles that represent influence of single factor.
  \end{itemize}
\end{frame}

\begin{frame}{Measuring ENSO Intensity}
  \begin{tikzpicture}[node distance = 1.5in, line width = 0.5mm]
    \node [process] (input) {
      \begin{varwidth}{1.2in}
        External forcing records and predictions
      \end{varwidth}};
    \node [process] (model) [right of=input] {
      \begin{varwidth}{1.2in}
        CESM1 and CESM2 climate models
      \end{varwidth}};
    \node [process] (output) [right of=model] {
      \begin{varwidth}{1.2in}
        Sea surface temperature (SST)
      \end{varwidth}};
    \node [process] (nino) [below of=input] {
      \begin{varwidth}{1.2in}
        Mean tropical Pacific temperature
      \end{varwidth}};
    \node [process] (variance) [right of=nino] {
      \begin{varwidth}{1.0in}
        Windowed variance: Mean variance over every 20 years.
      \end{varwidth}};
    \node [process] (intensity) [right of=variance] {
      \begin{varwidth}{1.2in}
        Mean and standard error
      \end{varwidth}};
    \draw [->] (input) to (model);
    \draw [->] (model) to (output);
    \draw [->] (output) to [out = 270, in = 90] (nino);
    \draw [->] (nino) to (variance);
    \draw [->] (variance) to (intensity);
  \end{tikzpicture}
\end{frame}

\begin{frame}{ENSO is Becoming Stronger}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
    \item Increase in ENSO intensity in both ensembles. (Exceeds 2 standard errors)
    \item Increase slows down in CESM1 and decreases in CESM2 after around 2050.
    \end{itemize}
    \column{0.5\textwidth}
    \begin{figure}
      \begin{tikzpicture}
        \node[anchor=south west,inner sep=0] at (0,0) {\includegraphics[width=\textwidth]{figures/ff_compare.pdf}};
        \coordinate (a) at (0.5, 1.5);
        \node [red] (b) at (1.0, 1.0) {
          \begin{varwidth}{0.6in}
            Increasing variance
          \end{varwidth}};
        \coordinate (c) at (3.0, 1.5) {};
        \draw[red, |-|] (a) to (c);

        \coordinate (d) at (3.2, 1.5);
        \node [red] (e) at (3.5, 1.0) {
          \begin{varwidth}{0.6in}
            Decreasing variance
          \end{varwidth}};
        \coordinate (f) at (4.3, 1.5) {};
        \draw[red, |-|] (d) to (f);
      \end{tikzpicture}
      \caption{ENSO intensity ensemble mean and standard error for CESM1 and CESM2}
      \label{ff_compare}
    \end{figure}
  \end{columns}
\end{frame}

\begin{frame}{Influence of Aerosols and Greenhouse Gasses}
  \begin{columns}
    \column{0.65\textwidth}
    \begin{itemize}
    \item Influence of each factor on ENSO amplitude.
    \item Increased variance due to greenhouse gas emissions.
    \item Somewhat increased variance from aerosol emissions, but not linear.
    \end{itemize}
    \alert{Takeaway:} Human activities are triggering predicted strengthening of ENSO.
    \column{0.5\textwidth}
    \myfig{cesm1_sf_4.pdf}{Influence of GHG, AER, and BMB forcing on ENSO amplitude in CESM1}{cesm1_sf_4}
  \end{columns}
\end{frame}

@@

% SPLIT UP CONCLUSIONS DISCUSSION

\begin{frame}{Discussion}
  \begin{itemize}
  \item Rising greenhouse gas levels strengthen ENSO cycle.
  \item Aerosol influence is nonlinear because aerosol levels are not purely increasing.
  \item Stronger ENSO may lead to greater temperature variability and extreme weather.
  \item External forcing affects lower frequency ENSO more.
  \end{itemize}
\end{frame}

@@

\begin{frame}{Acknowledgments}
  \begin{itemize}
  \item This material is based upon work supported by the National Center for Atmospheric Research, which is a major facility sponsored by the National Science Foundation under Cooperative Agreement No. 1852977.
  \item Thank you to my teacher, my family, and my mentor!
  \item Software used: R, ncdf4, zoo, dplyr, ggplot2, WaveletComp, reshape2, nco.
  \end{itemize}
\end{frame}

\begin{frame}{References}
  \bibliographystyle{apalike}
  \fontsize{4pt}{5}\selectfont
  \bibliography{references.bib}
\end{frame}

>>>>

\documentclass[aspectratio=1610]{beamer}

@@

\begin{frame}{Climate Change: Global Warming vs. Climate Variability}
  \begin{figure}
    \centering
    \includegraphics[width=0.9\textheight]{figures/intro_fig_3.pdf}
    \caption{Global mean land air temperature in GISSTEMP 4 dataset. \citep{gistemp2019giss} and \citep{lenssen2019improvements}}
    \label{fig:woohoo}
  \end{figure}
\end{frame}

\begin{frame}{Climate Forcing}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
    \item \alert{Forcing}: any external factor that affects climate.
      \begin{description}
      \item[\alert{GHG}] Greenhouse gasses
      \item[\alert{AER}] Aerosols (natural: volcanic ash, artificial: smoke)
      \item[\alert{BMB}] Biomass burning
      \item[\alert{LULC}] Land use/cover (deforestation, desertification)
      \end{description}
    \end{itemize}
    \column{0.6\textwidth}
    \myfig{greenhouse_Effect.jpg}{Factors that contribute to the greenhouse effect.}{this}
  \end{columns}
\end{frame}

\begin{frame}{El Niño, La Niña, and ENSO}
  \begin{figure}
    \begin{columns}
      \column{0.5\textwidth}
      \begin{itemize}
      \item ENSO = El Niño/Southern Oscillation
      \item El Niño is the warm phase, and La Niña is the cool phase.
      \item Warming and cooling of the Pacific Ocean.
      \item Affects human societies through temperature and rainfall. \citep{ropelewski1987global}
      \end{itemize}
      \caption{Comparison of SST anomaly between 1975 La Niña event and 1997 El Niño event in HadISST 1 dataset. \citep{rayner2003global}}
      \column{0.5\textwidth}
      \includegraphics[width = \textwidth]{figures/intro_fig.pdf}
    \end{columns}
  \end{figure}
\end{frame}

@@

\begin{frame}{Gap and Questions}
  \alert{Research Gap}
  \begin{itemize}
  \item Little research using a large ensemble to examine the effect of individual factors on ENSO.
  \item Considerable disagreement between studies on whether ENSO will strengthen or weaken due to global warming
  \end{itemize}
  \alert{Questions}
  \begin{enumerate}
  \item Do the CESM1 and CESM2 climate models predict increased or decreased ENSO intensity in the future?
  \item Is the predicted increase (or decrease) due to human activities?
  \end{enumerate}
\end{frame}

@@

\begin{frame}{Data: Community Earth System Model}
  \begin{itemize}
  \item Community Earth System Model (CESM) Versions 1 and 2 \citep{kay2015community} \citep{danabasoglu2020community}.
  \item Predicts climate over 21st century with global warming.
  \item Ensemble: collection of multiple simulations.
  \item Single forcing ensembles that represent influence of single factor.
  \item Software used: R, ncdf4, zoo, dplyr, ggplot2, WaveletComp, reshape2, nco.
  \end{itemize}
\end{frame}

\begin{frame}{Measuring ENSO Intensity}
  \begin{tikzpicture}[node distance = .37\textwidth, line width = 0.5mm]
    \node [process] (input) {
      \begin{varwidth}{1.2in}
        External forcing records and predictions
      \end{varwidth}};
    \node [process] (model) [right of=input] {
      \begin{varwidth}{1.2in}
        CESM1 and CESM2 climate models
      \end{varwidth}};
    \node [process] (output) [right of=model] {
      \begin{varwidth}{1.2in}
        Sea surface temperature (SST)
      \end{varwidth}};
    \node [process] (nino) [below of=input] {
      \begin{varwidth}{1.2in}
        Mean tropical Pacific temperature
      \end{varwidth}};
    \node [process] (variance) [right of=nino] {
      \begin{varwidth}{1.0in}
        Windowed variance: Mean variance over every 20 years.
      \end{varwidth}};
    \node [process] (intensity) [right of=variance] {
      \begin{varwidth}{1.2in}
        Mean and standard error
      \end{varwidth}};
    \draw [->] (input) to (model);
    \draw [->] (model) to (output);
    \draw [->] (output) to [out = 270, in = 90] (nino);
    \draw [->] (nino) to (variance);
    \draw [->] (variance) to (intensity);
  \end{tikzpicture}
\end{frame}

\begin{frame}{ENSO is Becoming Stronger}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
    \item Increase in ENSO intensity in both ensembles. (Exceeds 2 standard errors)
    \item Increase slows down in CESM1 and decreases in CESM2 after around 2050.
    \end{itemize}
    \column{0.5\textwidth}
    \begin{figure}
      \begin{tikzpicture}
        \node[anchor=south west,inner sep=0] at (0,0) {\includegraphics[width=\textwidth]{figures/ff_compare.pdf}};
        \coordinate (a) at (1.0, 1.5);
        \node [red] (b) at (1.3, 1.0) {
          \begin{varwidth}{0.6in}
            Increasing variance
          \end{varwidth}};
        \coordinate (c) at (3.8, 1.5) {};
        \draw[red, |-|] (a) to (c);

        \coordinate (d) at (4.0, 1.5);
        \node [red] (e) at (4.3, 1.0) {
          \begin{varwidth}{0.6in}
            Decreasing variance
          \end{varwidth}};
        \coordinate (f) at (5.0, 1.5) {};
        \draw[red, |-|] (d) to (f);
      \end{tikzpicture}
      \caption{ENSO intensity ensemble mean and standard error for CESM1 and CESM2}
      \label{ff_compare}
    \end{figure}
  \end{columns}
\end{frame}

\begin{frame}{Influence of Aerosols and Greenhouse Gasses}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
    \item Influence of each factor on ENSO amplitude.
    \item Increased variance due to greenhouse gas emissions.
    \item Somewhat increased variance from aerosol emissions, but not linear.
    \end{itemize}
    \alert{Takeaway:} Human activities are triggering predicted strengthening of ENSO.
    \column{0.5\textwidth}
    \myfig{cesm1_sf_4.pdf}{Influence of GHG, AER, and BMB forcing on ENSO amplitude in CESM1}{cesm1_sf_4}
  \end{columns}
\end{frame}

@@

\begin{frame}{Conclusion and Discussion}
  \begin{enumerate}
  \item Do the CESM1 and CESM2 climate models predict increased or decreased ENSO intensity in the future?
  \item Is the predicted increase (or decrease) due to human activities?
  \end{enumerate}
  \begin{enumerate}
  \item Yes, ENSO variance increases significantly over the 21$^{st}$ century
  \item Yes, the major factor contributing to this increase is greenhouse gasses, with aerosol emissions playing a secondary role.
  \end{enumerate}
\end{frame}

@@

\begin{frame}{Acknowledgments}
  \begin{itemize}
  \item This material is based upon work supported by the National Center for Atmospheric Research, which is a major facility sponsored by the National Science Foundation under Cooperative Agreement No. 1852977.
  \item Thank you to my teacher, my family, and my mentor!
  \end{itemize}
\end{frame}

\begin{frame}{References}
  \bibliographystyle{apalike}
  \fontsize{6pt}{5}\selectfont
  \bibliography{references.bib}
  \url{https://www.coolaustralia.org/the-greenhouse-effect-secondary}
\end{frame}
